---
title: "Ratio of Two Poisson Distributed Variables"
author: "Alexander Ondrus"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggfortify)
library(patchwork)

gen_single_plot_layer <- function(count_obs, confidence) {
  shape <- count_obs
  lower_bound <- qgamma(
    p = (1 - confidence) / 2,
    shape = count_obs,
    rate = 1
  ) %>%
    floor()
  upper_bound <- qgamma(
    p = (1 - confidence) / 2,
    shape = shape,
    rate = 1,
    lower.tail = FALSE
  ) %>%
    ceiling()

  p <- ggdistribution(
    dgamma,
    seq(lower_bound, upper_bound, 0.5),
    shape = shape,
    rate = 1,
    fill = "red"
  ) +
    labs(title = paste(
      paste("Shape =", shape),
      paste("Confidence =", confidence),
      sep = "\n"
    )) +
    scale_x_continuous(
      breaks = c(
        lower_bound,
        count_obs,
        upper_bound
      )
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )

  p <- ggdistribution(
    dgamma,
    seq(
      qgamma(p = 0.001, shape = count_obs, rate = 1),
      qgamma(p = 0.999, shape = count_obs, rate = 1),
      0.5
    ),
    shape = shape,
    rate = 1,
    p = p
  )
}
```

## Distribution of Poisson Average

We use Bayesian inference to construct credible intervals for the average parameter of a Poisson distributed variable. 

$$h(\lambda | x) = \frac{g(\lambda) \cdot f(x | \lambda)}{\int_0^\infty g(t) f(x | t) dt}$$ 

Assume a single observation with a count of $a$. Here I use an improper constant prior of 1. 

$$
\begin{matrix}
h(\lambda | a) &=& \frac{1 \cdot f(a | \lambda)}{\int_0^\infty 1\cdot f(a | t) dt} \\ 
&=& \frac{e^{-\lambda \frac{\lambda^a}{a!}}}{\int_0^\infty e^{-t}\frac{t^a}{a!} dt} \\
&=& \frac{e^{-\lambda} \lambda^a}{\Gamma(a + 1)} \\
&=& \frac{1^{a + 1}}{\Gamma(a+1)} \lambda^{(a+1)-1} e^{-1 \cdot \lambda}
\end{matrix}
$$

Thus, a single observation of a count of $a$ for a Poisson-distributed variable gives a $\mathrm{Gamma}(a + 1, 1)$ posterior distribution for $\lambda$.

### Examples

```{r plotting-egs, message=FALSE, warning=FALSE, fig.height=2}
gen_single_plot_layer(10, 0.9) + gen_single_plot_layer(100, 0.9) + gen_single_plot_layer(1000, 0.9)
```

## Ratio of Poisson Variables

Assuming $X \sim \mathrm{Poisson}(\lambda_1)$ and $Y \sim \mathrm{Poisson}(\lambda_2)$ we can construct credible intervals $a_1 \leq \lambda_1 \leq a_2$ and $b_1 \leq \lambda_2 \leq b_2$. The simplest interval for for $\frac{\lambda_1}{\lambda_2}$ is then $\frac{a_1}{b_2} \leq \frac{\lambda_1}{\lambda_2} \leq \frac{a_2}{b_1}$. We use the following function to construct this interval:

```{r ratio-function, echo=TRUE}
poisson_ratio <- function(count1, count2, confidence){
  prob_vector <- c((1 - confidence) / 2, 1 - ((1 - confidence)/2))
  count1_bounds <- qgamma(
    prob_vector,
    shape = count1,
    rate = 1
  )
  count2_bounds <- qgamma(
    prob_vector,
    shape = count2,
    rate = 1
  )
  
  return_list = 
    list(
    "Numerator Bounds" = count1_bounds,
    "Denominator Bounds" = count2_bounds,
    "Ratio Bounds" = c(
      min(count1_bounds) / max(count2_bounds),
      max(count1_bounds) / min(count2_bounds)
    )
  )
  
  return(return_list)
}
```

### Example Calculation

Suppose that a given process observed a count of 1354 last year and after going through a dramatic change and observed a count of 609 this year. Construct a 90% credible interview for the ratio of true average observations from one year to the next?

First, we visualize the credible intervals for each average separately:

```{r visualize-egs, message=FALSE, warning=FALSE, fig.height=2}
gen_single_plot_layer(1354, 0.9) + gen_single_plot_layer(609, 0.9)
```

Using the function defined above we see that the 90% credible interval for the ratio of the average to be `r round(poisson_ratio(count1=609, count2=1354, 0.9)[["Ratio Bounds"]][1], 3)` $\leq$ `r round(609 / 1354, 3)` $\leq$ `r round(poisson_ratio(count1=609, count2=1354, 0.9)[["Ratio Bounds"]][2], 3)`.
